---
title: "Retail store Casestudy"
author: "Jashwanth Neeli"
date: "16/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# BUSINESS PROBLEM: 



#### A Retail store is required to analyze the day-to-day transactions and keep a 
#### track of its customers spread across various locations along with their 
#### purchases/returns across various categories. 
#### Create an RMarkdown report and display the below calculated metrics, 
#### reports and inferences


### DATA AVAILABLE -

#### Customer: Customers information including demographics 
#### Transaction: Transactions of customers 
#### Product Hierarchy: Product information (category, sub category etc...)



#### 1. Merge the datasets Customers, Product Hierarchy and Transactions as 
#### Customer_Final. Ensure to keep all customers who have done 
#### transactions with us and select the join type accordingly. 


#### a. Use the base merge() 


```{r}
setwd("C:/R_Aug/case_study1_Retail")
f1 = read.csv("Customer.csv")
f2 = read.csv("prod_cat_info.csv")
f3 = read.csv("Transactions.csv")
```


#a: using the base merge:

#####Method1: using (core R method)

#####Common column that can be used to perform the join is customer_Id in Customer table and cust_id in 
##Transactions table and prod_cat table can be combined using prod_cat_code and prod_sub_cat_code 
#from prod_cat table. (Inner JOIN is Used here as i only want common records.)



```{r}
merging <- merge(x=f1, y=f3, by.x="customer_Id", by.y="cust_id", all= FALSE)


Customer_Final <- merge(x=merging, y=f2, by.x= c("prod_cat_code","prod_subcat_code"), by.y= c("prod_cat_code", "prod_sub_cat_code"), all = FALSE)

```

##### Checking NA Value In  The Dataset

```{r}

sum(is.na(Customer_Final))

```
##### Droping NA Value as Number of NA Value is very less

``` {r} 

require(tidyr)
Customer_Final <- tidyr::drop_na(Customer_Final)

Customer_Final <- inner_join(dplyr_merge, f2, by=c("prod_cat_code"="prod_cat_code", "prod_subcat_code"="prod_sub_cat_code"))

```

#### b. Dplyr merge functions 

```{r}


require(dplyr)
dplyr_merge <- inner_join(f1, f3, by=c("customer_Id"="cust_id"))

```

```{r}

df<- Customer_Final
dim(df)

```
#2.prepare a summary report for the merged data set.

#a. Get the column names and their corresponding data types.

```{r}

#2.prepare a summary report for the merged data set.

#a. Get the column names and their corresponding data types.

    col_names <- data.frame(colnames(df))
    data.frame(sapply(df, class))

```


###b.  top/bottom 10 observations.

```{r}

    
    View(head(df,10))

    View(tail(df,10))

```


#c.  "Five-number summary" for continuous variables(min, Q1, Median, Q3 and max)

```{r}

    isnum <- sapply(df, is.numeric)
    df_num <- df[,isnum]
    summary(df_num)

```

#d.  frequency tables for all the categorical variables.

```{r}
    
FREQ_STORETYPE <- Customer_Final %>% dplyr::group_by(Store_type) %>% dplyr::summarise(FREQ = 
                                                                                        n()) %>% dplyr::arrange( desc(FREQ))
FREQ_STORETYPE
FREQ_PRODCAT <- Customer_Final %>% dplyr::group_by(prod_cat) %>% dplyr::summarise(FREQ = 
                                                                                    n()) %>% dplyr::arrange( desc(FREQ))
FREQ_PRODCAT
FREQ_PRODSUB <- Customer_Final %>% dplyr::group_by(prod_subcat) %>% dplyr::summarise(FREQ =
                                                                                       n()) %>% dplyr::arrange( desc(FREQ))
FREQ_PRODSUB


```

#3. Generate histograms for all continuous variables and frequency bars for categorical variables.
    
#histograms for all continuous variables

```{r}
    
require(ggplot2)  

ggplot2::ggplot(data = df_num) + aes(x= prod_cat_code) + geom_histogram(bins = 6, color = "red", fill = "skyblue2")



ggplot2::ggplot(data = df_num) + aes(x= Tax) + geom_histogram(bins = 10, color = "red", fill = "skyblue2") + ggtitle("Tax_Amount")



ggplot2::ggplot(data = df_num) + aes(x= total_amt) + geom_histogram(bins = 10, color = "red", fill = "skyblue2") + ggtitle("Tax_Amount")


```


```{r}

#FREQUENCY BARS FOR CATEGORICAL VARIABLE.

#frequency bar for product category.
prod_cat_freq <- df_char%>% group_by(prod_cat) %>% summarise(freq = n())

ggplot2::ggplot(data = prod_cat_freq) + aes(x= prod_cat, y= freq) + geom_bar(stat = "identity", fill= "red", color = "black" ) +ggtitle("count of each prod_cat")


#frequency bar for product sub category.
prod_subcat_freq <- df_char%>% group_by(prod_subcat) %>% summarise(freq = n())

ggplot2::ggplot(data = prod_subcat_freq) + aes(x= prod_subcat, y= freq) + geom_bar(stat = "identity", fill= "red", color = "black" ) + ggtitle("count of each product sub_cat")


#frequency bar for store type category.
Store_type_freq <- df_char%>% group_by(Store_type) %>% summarise(freq = n())

ggplot2::ggplot(data = Store_type_freq) + aes(x= Store_type, y= freq) + geom_bar(stat = "identity", fill= "red", color = "black" ) + ggtitle("count of each storetype")


#frequency bar for Gender category.
Gender_freq <- df_char%>% group_by(Gender) %>% summarise(freq = n())

ggplot2::ggplot(data = Gender_freq) + aes(x= Gender, y= freq) + geom_bar(stat = "identity", fill= "red", color = "black" ) + ggtitle("count of Gender")



```

#4. Calculate the following information using the merged dataset:

#a. Time period of the available transaction data.

```{r}

require(lubridate)

## Loading required package: lubridate
## 
## Attaching package: 'lubridate'
## The following objects are masked from 'package:base':
## 
##     date, intersect, setdiff, union

df$TRAN_DATE_FORMAT<- lubridate::dmy(df$tran_date)

A <- min(df$TRAN_DATE_FORMAT)
B <- max(df$TRAN_DATE_FORMAT)


(B-A)


```

#b.  count of transactions where the total amount of transaction was negative.

```{r}

#b.  count of transactions where the total amount of transaction was negative.

neg_trans <- df%>% dplyr::filter(total_amt<0)

neg_trans %>% dplyr::distinct(neg_trans$customer_Id) %>%summarise(trans_cnt=n())

```


#5. Analyze which product categories are more popular among females vs male customers.

```{r}

#5. Analyze which product categories are more popular among females vs male customers.

Customer_male <- df[(df$Gender == "M"),]
popular_prodM <- Customer_male %>% dplyr::group_by(Gender,prod_cat)%>%dplyr::summarise(ttl_cnt=sum(Qty))%>% dplyr::arrange(desc(ttl_cnt))
popular_prodM

Customer_Female <- df[(df$Gender == "F"),]
popular_prodF <- Customer_Female %>% dplyr::group_by(Gender,prod_cat)%>%dplyr::summarise(ttl_cnt=sum(Qty))%>% dplyr::arrange(desc(ttl_cnt))
popular_prodF

pop_vs_Mandf <- cbind(popular_prodM, popular_prodF)
colnames(pop_vs_Mandf) <- c("GenderM", "Prod_catM", "ttl_qty", "GenderF", "prod_catF", "TTl_qtyF")
pop_vs_Mandf


```

#6. Which city code has the maximum customers and what was the percentage of customers from that city?

```{r}

count_citycode <- df%>% dplyr::group_by(city_code)%>% dplyr::summarise(cnt_cust=n(), percentage = round(n()*100/length(df$city_code),2))%>%dplyr::arrange(desc(percentage))

View(count_citycode)

```


#7. Which store type sells the maximum products by value and by quantity?

```{r}

df %>% dplyr::group_by(Store_type)%>% summarise(sum_qty=sum(Qty),sum_val=sum(total_amt))


```


#8. What was the total amount earned from the “Electronics” and “Clothing” categories from Flagship Stores?

```{r}

Cust_flagship <- df[(df$Store_type=="Flagship store"),]
category <- Cust_flagship[(Cust_flagship$prod_cat=="Electronics")|(Cust_flagship$prod_cat=="Clothing"),]
ttlamt_earned <- category %>% dplyr::group_by(Store_type,prod_cat)%>% summarise(ttl_amt=sum(total_amt))
ttlamt_earned


```


#9. What was the total amount earned from “Male” customers under the “Electronics” category?

```{r}

Cust_male <- df[df$Gender=="M",]
male_Electronics <- Cust_male[Cust_male$prod_cat=="Electronics",]
male_Electronics%>% dplyr::group_by(Gender,prod_cat)%>%summarise(ttl_earn=sum(total_amt))

```


#10. How many customers have more than 10 unique transactions, after removing all transactions which have any negative amounts?

```{r}

nonnegative_trans <- df[df$total_amt >0,]
cust_trans <- nonnegative_trans %>% dplyr::group_by(customer_Id)%>%summarise(cnt= n())
unique_cust_trans <- cust_trans[cust_trans$cnt>10,]
unique_cust_trans


```

#11.  For all customers aged between 25 - 35, find out:

#a. What was the total amount spent for for “Electronics” and “Books” product categories?

```{r}

require("lubridate")
DATE <- lubridate::dmy(df$DOB)
tran_date <- lubridate::dmy(df$tran_date)

df$DIFFERENCE <- df$tran_date-DATE

AGE25_35_DATA <- df[(df$DIFFERENCE > 25*365.25) & (df$DIFFERENCE < 35*365.25), ]
AGE25_35_ELCBOOK <- AGE25_35_DATA[ (AGE25_35_DATA$prod_cat == "Electronics") | (AGE25_35_DATA$prod_cat =="Books"),]

AGEBOOK_ELC <- AGE25_35_ELCBOOK %>% dplyr::group_by(prod_cat) %>% summarise(Total_Sum = sum(total_amt)) %>%dplyr::arrange(desc(prod_cat))

AGEBOOK_ELC


```


#b. What was the total amount spent by these customers between 1st Jan, 2014 to 1st Mar, 2014?

```{r}



DATEBETWEEN_JANMARCH <- AGE25_35_DATA[(AGE25_35_DATA$tran_date >= "2014-01-01") & (AGE25_35_DATA$tran_date <= "2014-03-01"),]

AMOUNT_SPENT_BETWEEN <- DATEBETWEEN_JANMARCH  %>% dplyr::summarise(REVENUE = sum(total_amt))

AMOUNT_SPENT_BETWEEN

```

